---
title: "Women in German Local Politics"
output:
  flexdashboard::flex_dashboard:
  runtime: shiny
---

```{r setup, include=FALSE}
# required packages
p_needed <-
  c(
    "viridis",
    "knitr",
    "ggplot2",
    "dplyr",
    "tidyr",
    "haven",
    "readxl",
    "viridis",
    "sf",
    "stringr",
    "leaflet",
    "rjson",
    "jsonlite",
    "RCurl",
    "leaflet.extras",
    "RColorBrewer",
    "htmltools",
    "htmlwidgets",
    "shiny",
    "DT",
    "flexdashboard",
    "plotly"
  )

# installed packages
packages <- rownames(installed.packages())

# missing packages
p_to_install <- p_needed[!(p_needed %in% packages)]

# install missing packages
if (length(p_to_install) > 0) {
  install.packages(p_to_install, repos = "http://cran.rstudio.com")
}

# load packages
sapply(p_needed, require, character.only = TRUE)

# #for shiny.io
# library(viridis)
# library(knitr)
# library(ggplot2)
# library(dplyr)
# library(tidyr)
# library(haven)
# library(readxl)
# library(viridis)
# library(sf)
# library(stringr)
# library(leaflet)
# library(rjson)
# library(jsonlite)
# library(RCurl)
# library(leaflet.extras)
# library(RColorBrewer)
# library(htmltools)
# library(htmlwidgets)
# library(shiny)
# library(DT)
# library(flexdashboard)
# library(plotly)

#working directory
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

#load necessary data files and manipulate them

load("kommunalwahlen_2000_2022.RData") #municipal election data

eige_data <- read.csv("eige_data.csv") #european data
eige_data <- pivot_wider(eige_data, names_from = POSITION, values_from = Mayor)
names(eige_data) <- c("Year", "Country", "Share of Female Mayors", "Share of Female Councillors")

#spatial data
my_spdf <- st_read(
  dsn = "tempdir",
  layer = "torn"
)

#manipulate data for data table
alle_wahlen_dt <- alle_wahlen %>%
  dplyr::mutate(
    Bundesland = bundesland,
    `Election Date` = datum_kommunalwahl,
    Municipality = gemeinde_name,
    `Total Council Seats` = sitze_gesamt,
    `Female Councillors` = sitze_frauen,
    `Share of Female Councillors` = sitze_frauen/sitze_gesamt*100,
    `CDU-CSU Councillors` = sitze_gesamt_union,
    `Share of CDU-CSU Councillors` = sitze_gesamt_union/sitze_gesamt*100,
    `SPD Councillors` = sitze_gesamt_spd,
    `Share of SPD Councillors` = sitze_gesamt_spd/sitze_gesamt*100,
    `Green Councillors` = sitze_gesamt_gruene,
    `Share of Green Councillors` = sitze_gesamt_gruene/sitze_gesamt*100,
    `FDP Councillors` = sitze_gesamt_fdp,
    `Share of FDP Councillors` = sitze_gesamt_fdp/sitze_gesamt*100,
    `Left Councillors` = sitze_gesamt_linke,
    `Share of Left Councillors` = sitze_gesamt_linke/sitze_gesamt*100,
    `AfD Councillors` = sitze_gesamt_afd,
    `Share of AfD Councillors` = sitze_gesamt_afd/sitze_gesamt*100,
    `Female CDU-CSU Councillors` = sitze_frauen_union,
    `Share of Female CDU-CSU Councillors` = sitze_frauen_union/sitze_gesamt_union*100,
    `Female SPD Councillors` = sitze_frauen_spd,
    `Share of Female SPD Councillors` = sitze_frauen_spd/sitze_gesamt_spd*100,
    `Female Green Councillors` = sitze_frauen_gruene,
    `Share of Female Green Councillors` = sitze_frauen_gruene/sitze_gesamt_gruene*100,
    `Female FDP Councillors` = sitze_frauen_fdp,
    `Share of Female FDP Councillors` = sitze_frauen_fdp/sitze_gesamt_fdp*100,
    `Female Left Councillors` = sitze_frauen_linke,
    `Share of Female Left Councillors` = sitze_frauen_linke/sitze_gesamt_linke*100,
    `Female AfD Councillors` = sitze_frauen_afd,
    `Share of Female AfD Councillors` = sitze_frauen_afd/sitze_gesamt_afd*100,
    `Female Mayor` = as.factor(ifelse(bürgermeisterin == "1", "yes", ifelse(bürgermeisterin == "0","no", NA))),
    `Party of Mayor` = partei_bürgermeister,
    `Date of Mayoral Election` = datum_bürgermeisterwahl,
    `Total Number of Candidates` = bewerber_gesamt,
    `Female Number of Candidates` = bewerber_frauen,
    `Share of Female Candidates` = bewerber_frauen/bewerber_gesamt*100
  ) %>%
  dplyr::select(Bundesland,
    `Election Date`,
    Municipality,
    `Total Council Seats`,
    `Female Councillors`,
    `Share of Female Councillors`,
    `CDU-CSU Councillors`,
    `Share of CDU-CSU Councillors`,
    `SPD Councillors`,
    `Share of SPD Councillors`,
    `Green Councillors`,
    `Share of Green Councillors`,
    `FDP Councillors`,
    `Share of FDP Councillors`,
    `Left Councillors`,
    `Share of Left Councillors`,
    `AfD Councillors`,
    `Share of AfD Councillors`,
    `Female CDU-CSU Councillors`,
    `Share of Female CDU-CSU Councillors`,
    `Female SPD Councillors`,
    `Share of Female SPD Councillors`,
    `Female Green Councillors`,
    `Share of Female Green Councillors`,
    `Female FDP Councillors`,
    `Share of Female FDP Councillors`,
    `Female Left Councillors`,
    `Share of Female Left Councillors`,
    `Female AfD Councillors`,
    `Share of Female AfD Councillors`,
    `Female Mayor`,
    `Party of Mayor`,
    `Date of Mayoral Election`,
    `Total Number of Candidates`,
    `Female Number of Candidates`,
    `Share of Female Candidates`)

#manipulate data for correlation plots
alle_wahlen_corr <- alle_wahlen %>%
  dplyr::mutate(
    `Election Date` = datum_kommunalwahl,
    `Total Seats` = sitze_gesamt,
    `Share of Female Councillors` = sitze_frauen / sitze_gesamt * 100,
    `Share of Female Candidates` = bewerber_frauen / bewerber_gesamt *
      100,
    `Total Candidates` = bewerber_gesamt,
    `Seat Share CDU` = sitze_gesamt_union / sitze_gesamt * 100,
    `Seat Share SPD` = sitze_gesamt_spd / sitze_gesamt * 100,
    `Seat Share Green Party` = sitze_gesamt_gruene / sitze_gesamt *
      100,
    `Seat Share FDP` = sitze_gesamt_fdp / sitze_gesamt * 100,
    `Seat Share Left Party` = sitze_gesamt_linke / sitze_gesamt * 100,
    `Seat Share AfD` = sitze_gesamt_afd / sitze_gesamt * 100,
    `Female Mayor` = as.factor(ifelse(bürgermeisterin == "1", "yes", ifelse(bürgermeisterin == "0", "no", NA))),
    `Female Minister` = as.factor(ifelse(ministerpräsidentin == "1", "yes", ifelse(ministerpräsidentin == "0", "no", NA))),
    Bundesland = as.factor(bundesland)
  ) %>%
  dplyr::select(`Election Date`, `Total Seats`, `Share of Female Councillors`, `Share of Female Candidates`, `Total Candidates`, `Seat Share CDU`, `Seat Share SPD`, `Seat Share Green Party`, `Seat Share FDP`, `Seat Share Left Party`, `Seat Share AfD`, `Female Mayor`, `Female Minister`, Bundesland)
```




Councillors in Municipal Parliaments
=======================================================================


Column {data-width=150}
-----------------------------------------------------------------------

### Number of Municipal Councils in Germany

```{r two, cache=TRUE}
num_councils <- 11087
flexdashboard::valueBox(prettyNum(num_councils, big.mark = ","),
         caption = "Municipal Councils in Germany")
```

### Number of Female Municipal Councillors in Germany

```{r three, cache=TRUE}
num_female_councillors <- 39434
flexdashboard::valueBox(prettyNum(num_female_councillors, big.mark = ","),
         caption = "Female Municipal Councillors in Germany")
```

### Number of Municipal Councillors in Germany

```{r four, cache=TRUE}
num_councillors <- 165353
flexdashboard::valueBox(prettyNum(num_councillors, big.mark = ","),
         caption = "Municipal Councillors in Germany")
```



Column {data-width=150}
-----------------------------------------------------------------------

### Councils with at least 50% Women

```{r five, cache=TRUE}
councils_total <- alle_wahlen %>%
  select(sitze_gesamt, sitze_frauen) %>%
  filter(!is.na(sitze_gesamt)) %>%
  filter(!is.na(sitze_frauen)) %>%
  count()

councils_plus_fifty <- alle_wahlen %>%
  select(sitze_gesamt, sitze_frauen) %>%
  filter(!is.na(sitze_gesamt)) %>%
  filter(!is.na(sitze_frauen)) %>%
  mutate(percent = sitze_frauen/sitze_gesamt) %>%
  filter(percent >= 0.5) %>%
  count()

equal_councils <- councils_plus_fifty/councils_total*100
equal_councils <- as.numeric(equal_councils)

flexdashboard::gauge(value = equal_councils,
      min = 0,
      max = 100,
      symbol = "%")
```


### Councils with no Women

```{r six, cache=TRUE}
councils_total <- alle_wahlen %>%
  select(sitze_gesamt, sitze_frauen) %>%
  filter(!is.na(sitze_gesamt)) %>%
  filter(!is.na(sitze_frauen)) %>%
  count()

councils_zero <- alle_wahlen %>%
  select(sitze_gesamt, sitze_frauen) %>%
  filter(!is.na(sitze_gesamt)) %>%
  filter(!is.na(sitze_frauen)) %>%
  filter(sitze_frauen == 0) %>%
  count()

zero_councils <- councils_zero/councils_total*100
zero_councils <- as.numeric(zero_councils)

flexdashboard::gauge(value = zero_councils,
      min = 0,
      max = 100,
      symbol = "%")
```


### Share of Female Councillors

```{r seven, cache=TRUE}
female_share <- 23.8

flexdashboard::gauge(value = female_share,
      min = 0,
      max = 100,
      symbol = "%")
```


Country Comparisons
====================================================================

Column {data-width=300 .sidebar}
-----------------------------------------------------------------------
```{r eight}
country_choices <- eige_data %>% distinct(Country)

selectInput(
  "country_selector_councillors",
  label = "Select Countries for Comparison",
  choices = country_choices,
  multiple = TRUE,
  selected = "Germany"
)

filtered_eige_data <-
  reactive(eige_data %>% dplyr::filter(Country %in% input$country_selector_councillors))
```


Column {.tabset}
-----------------------------------------------------------------------

### Female Councillors over Time

```{r nine}
renderPlotly({
  line <- filtered_eige_data() %>%
    ggplot(aes(
      x = Year,
      y = `Share of Female Councillors`,
      group = Country,
      color = Country
    )) +
    geom_line() +
    theme_bw() + expand_limits(y = 0)
  
  ggplotly(line)
})
```

> Source: European Institute for Gender Equality (2024).

### Female Mayors over Time

```{r ten}
renderPlotly({
  line <- filtered_eige_data() %>%
    ggplot(aes(
      x = Year,
      y = `Share of Female Mayors`,
      group = Country,
      color = Country
    )) +
    geom_line() +
    theme_bw() + expand_limits(y = 0)
  
  ggplotly(line)
})
```

> Source: European Institute for Gender Equality (2024).


Correlation Explorer
=======================================================================

Column {data-width=300 .sidebar}
-----------------------------------------------------------------------
```{r eleven, cache=TRUE}

varSelectInput("corr_x",
               label = "Select variable to display on x-axis",
               data = alle_wahlen_corr[, c(1:11)])

varSelectInput("corr_y",
               label = "Select variable to display on y-axis",
               data = alle_wahlen_corr[, c(2:11)])

varSelectInput("corr_col",
               label = "Select variable to color",
               data = alle_wahlen_corr[, c(12:14)])

selectInput(
  "corr_trend",
  label = "Select to show trends",
  choices = c("None", "Linear Trend", "Linear Trend with Confidence Intervals")
)

```


Column
-----------------------------------------------------------------------

### Plot Two Variables Against One Another

```{r twelve}
renderPlotly({
  plot_minimal <- alle_wahlen_corr %>%
    ggplot(aes(
      x = !!input$corr_x,
      y = !!input$corr_y,
      col = !!input$corr_col
    )) +
    geom_point(alpha = .6) +
    theme_bw()
  
  plot_lm  <- alle_wahlen_corr %>%
    ggplot(aes(
      x = !!input$corr_x,
      y = !!input$corr_y,
      col = !!input$corr_col
    )) +
    geom_point(alpha = .6) +
    geom_smooth(method = lm ,
                color = "red",
                se = FALSE) +
    theme_bw()
  
  plot_lm_ci  <- alle_wahlen_corr %>%
    ggplot(aes(
      x = !!input$corr_x,
      y = !!input$corr_y,
      col = !!input$corr_col
    )) +
    geom_point(alpha = .6) +
    geom_smooth(method = lm ,
                color = "red",
                se = TRUE) +
    theme_bw()
  
  if (input$corr_trend == "None") {
    ggplotly(plot_minimal)
  }
  
  else if (input$corr_trend == "Linear Trend") {
    ggplotly(plot_lm)
  }
  
  else {
    ggplotly(plot_lm_ci)
  }
})
```

> Write some more info about this plot.


Map
=========================================================================

Column {data-width=300 .sidebar}
-----------------------------------------------------------------------
```{r thirteen}
selectInput(
  "map_variable",
  label = "Select variable to display",
  choices = c("Female Councillors", "Female Candidates", "Female Mayors")
)
```

Column
------------------------------------------------------------------------

### Map

```{r fourteen}
####Plot 1: Female Councillors

#create a color palette with handmade bins
mybins <- c(0, 10 , 20 , 30, 40, 50, 100)
mypalette <-
  colorBin(
    palette = "YlOrBr",
    domain = my_spdf$antl_frn_s,
    na.color = "transparent",
    bins = mybins
  )

#prepare the text for tooltips
mytext <- paste(
  "Municipality: ",
  my_spdf$NAME_4,
  "<br/>",
  "Date of Election: ",
  my_spdf$dtm_k,
  "<br/>",
  "Total Seats: ",
  my_spdf$stz_g,
  "<br/>",
  "Female Councillors: ",
  my_spdf$stz_f,
  "<br/>",
  "Share of Female Councillors: ",
  round(my_spdf$antl_frn_s, 2),
  "%",
  sep = ""
) %>%
  lapply(htmltools::HTML)

#final map
map_female_councillors <- leaflet(my_spdf) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~ mypalette(antl_frn_s),
    stroke = TRUE,
    fillOpacity = 0.9,
    color = "white",
    weight = 0.3,
    label = mytext,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "13px",
      direction = "auto"
    )
  ) %>%
  addLegend(
    pal = mypalette,
    values = ~ antl_frn_s,
    opacity = 0.9,
    title = "Share of female councillors (%)",
    position = "bottomleft"
  ) %>%
  addProviderTiles("Esri.WorldGrayCanvas") %>%
  leaflet.extras::addSearchOSM(options = searchOptions(collapsed = TRUE))

####Plot 2: Female Mayors

#create a color palette with handmade bins
mybins <- c(0, 1)
mypalette <-
  colorFactor(palette = "YlOrBr",
              domain = my_spdf$brgrm,
              na.color = "transparent")

#prepare the text for tooltips
mytext <- paste(
  "Municipality: ",
  my_spdf$NAME_4,
  "<br/>",
  "Female Mayor: ",
  my_spdf$brgrm,
  "<br/>",
  sep = ""
) %>%
  lapply(htmltools::HTML)

#final map
map_female_mayors <- leaflet(my_spdf) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~ mypalette(brgrm),
    stroke = TRUE,
    fillOpacity = 0.9,
    color = "white",
    weight = 0.3,
    label = mytext,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "13px",
      direction = "auto"
    )
  ) %>%
  addLegend(
    pal = mypalette,
    values = ~ brgrm,
    opacity = 0.9,
    title = "Female Mayor",
    position = "bottomleft"
  ) %>%
  addProviderTiles("Esri.WorldGrayCanvas") %>%
  leaflet.extras::addSearchOSM(options = searchOptions(collapsed = TRUE))


####Plot 3: Female Candidates

#create a color palette with handmade bins
mybins <- c(0, 10 , 20 , 30, 40, 50, 100)
mypalette <-
  colorBin(
    palette = "YlOrBr",
    domain = my_spdf$antl_frn_b,
    na.color = "transparent",
    bins = mybins
  )

#prepare the text for tooltips
mytext <- paste(
  "Municipality: ",
  my_spdf$NAME_4,
  "<br/>",
  "Date of Election: ",
  my_spdf$dtm_k,
  "<br/>",
  "Total Candidates: ",
  my_spdf$bwrbr_g,
  "<br/>",
  "Female Candidates: ",
  my_spdf$bwrbr_f,
  "<br/>",
  "Share of Female Candidates: ",
  round(my_spdf$antl_frn_b, 2),
  "%",
  sep = ""
) %>%
  lapply(htmltools::HTML)

#final map
map_female_candidates <- leaflet(my_spdf) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~ mypalette(antl_frn_b),
    stroke = TRUE,
    fillOpacity = 0.9,
    color = "white",
    weight = 0.3,
    label = mytext,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "13px",
      direction = "auto"
    )
  ) %>%
  addLegend(
    pal = mypalette,
    values = ~ antl_frn_b,
    opacity = 0.9,
    title = "Share of female candidates (%)",
    position = "bottomleft"
  ) %>%
  addProviderTiles("Esri.WorldGrayCanvas")



#######Plot them
renderLeaflet({
  if (input$map_variable == "Female Councillors") {
    map_female_councillors
  }
  
  else if (input$map_variable == "Female Candidates") {
    map_female_candidates
  }
  
  else if (input$map_variable == "Female Mayors") {
    map_female_mayors
  }
  
  else {
    map_female_councillors
  }
})

```


Raw Data
=======================================================================

Column {data-width=300 .sidebar}
-----------------------------------------------------------------------
```{r fifteen}
checkboxGroupInput(
  "data_selector",
  label = "Select variables for data table",
  choices = names(alle_wahlen_dt),
  selected = names(alle_wahlen_dt)
)
```

Column
-----------------------------------------------------------------------

### Data Table

```{r sixteen}
renderDataTable({
  datatable(
    alle_wahlen_dt %>% dplyr::select(input$data_selector),
    rownames = FALSE,
    extensions = "Buttons",
    filter = 'top',
    options = list(
      dom = "Bfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print")
    )
  )
})
```
